Versioned Queries over RDF Archives:\\All You Need is SPARQL?

### 2020-09-09 paper.tex
submitted version

A key aspect of the Web is its dynamic nature, where documents are frequently updated, deleted and added. Likewise when we speak of the Semantic Web, it is important to consider that sources may be dynamic and RDF datasets are subject to change~\cite{KaferAUOH13}. It is in this context that various works have looked at versioning in the context of RDF/SPARQL~\cite{VolkelG06,TappoletB09,Grandi10,GraubeHU14,KhuranaD16}, with recent works proposing \textit{RDF archives}~\cite{FernandezPU15,Cerdeira-PenaFF16,BahriLA18,FernandezUPK19,TaelmanSHMV19} that manage RDF graphs and their historical changes, allowing for querying across different versions of the graph. Within these works, a variety of specialised indexing techniques~\cite{Cerdeira-PenaFF16,BahriLA18,TaelmanSHMV19}, query languages~\cite{TappoletB09} and benchmarks~\cite{KotsevMPEFK16,FernandezUPK19} have been proposed, developed and evaluated. While these represent important advances, many such works propose custom SPARQL extensions, engines, etc., creating an barrier for adoption.

@@

In fact, the types of versioned queries proposed in the literature~\cite{FernandezPU15} can be supported using off-the-shelf SPARQL engines that already enjoy years of development, optimisation, and deployment. SPARQL named graphs can, for exampe, be used to track different versions of individual graphs. However, as Fernandez at al.~\cite{FernandezUPK19} note, the approach of using pure SPARQL would ``\textit{typically render rather inefficient SPARQL queries}''. This raises a question: how inefficient will such queries be? If a pure SPARQL solution could be found with reasonable performance, existing SPARQL infrastructure could be leveraged for RDF archives.

@@

More recently the concept of RDF archives have been gaining more and more attention. Fernandez et al.~\cite{FernandezPU15} provide a general survey and introduction to the theme, discussing the types of queries that one might wish to answer in the context of an RDF archive. Works by Cerdeira{-}Pena et al.~\cite{Cerdeira-PenaFF16}, Zaniolo et al.~\cite{ZanioloGACG18} and Taelman et al.~\cite{TaelmanSHMV19} propose ways to compress and index RDF archives, while Khurana and Deshpande~\cite{KhuranaD16} propose indexes more generally for historical graph data. Bahri et al.~\cite{BahriLA18} explore the benefits of using Apache Spark to manage RDF archives in a distributed setting. In order to evaluate these emerging techniques, a number of benchmarks have been proposed, including the BEnchmark of RDF
ARchives (BEAR)~\cite{FernandezUPK19}, and the Semantic Publishing Benchmark (SPB)~\cite{Papakonstantinou18}.

@@

We now reflect back on our research questions: (Q1) In terms of indexing space and time, incremental builds with an initial base version (\cbpd) are best. (Q2) In general the uncompressed IC representation (\ic) offers the best query runtimes, but interval TB (\tb) is quite competitive, and even outperforms IC for delta-version queries. (Q3) Rather than there being an overall winner, we note a time--space tradeoff where the larger representations offer faster queries and smaller representations slower queries. However, interval TB (\tb) arguably strikes the best balance if comparing both space and time.

For future work, it would be of interest to run experiments for other SPARQL engines and other RDF archive benchmarks. Also it would be interesting to run more diverse types of versioned queries, such as delta versions with large gaps, queries returning versions as solutions, etc. There are also open questions relating to more optimal/concise query rewritings, and support for paths.\footnote{Online material: \url{https://github.com/HunterNacho/sparql-versioning}}

>>>>

A key aspect of the Web is its dynamic nature, where documents are frequently updated, deleted and added. Likewise when we speak of the Semantic Web, it is important to consider that sources may be dynamic and RDF datasets are subject to change~\cite{KaferAUOH13}. It is in this context that various works have looked at versioning in the context of RDF/SPARQL~\cite{VolkelG06,TappoletB09,Grandi10,GraubeHU14,KhuranaD16}, with recent works proposing \textit{RDF archives}~\cite{FernandezPU15,Cerdeira-PenaFF16,BahriLA18,FernandezUPK19,TaelmanSHMV19} that manage RDF graphs and their historical changes, allowing for querying across different versions of the graph. Within these works, a variety of specialised indexing techniques~\cite{Cerdeira-PenaFF16,BahriLA18,TaelmanSHMV19}, query languages~\cite{TappoletB09} and benchmarks~\cite{KotsevMPEFK16,FernandezUPK19} have been proposed, developed and evaluated. While these represent important advances, many such works propose custom SPARQL extensions, engines, etc., creating a barrier for adoption.

@@

In fact, the types of versioned queries proposed in the literature~\cite{FernandezPU15} can be supported using off-the-shelf SPARQL engines that already enjoy years of development, optimisation, and deployment. SPARQL named graphs can, for example, be used to track different versions of individual graphs. However, as Fernandez at al.~\cite{FernandezUPK19} note, the approach of using pure SPARQL would ``\textit{typically render rather inefficient SPARQL queries}''. This raises a question: how inefficient will such queries be? If a pure SPARQL solution could be found with reasonable performance, existing SPARQL infrastructure could be leveraged for RDF archives.

@@

More recently the concept of RDF archives has been gaining more and more attention. Fernandez et al.~\cite{FernandezPU15} provide a general survey and introduction to the theme, discussing the types of queries that one might wish to answer in the context of an RDF archive. Works by Cerdeira{-}Pena et al.~\cite{Cerdeira-PenaFF16}, Zaniolo et al.~\cite{ZanioloGACG18} and Taelman et al.~\cite{TaelmanSHMV19} propose ways to compress and index RDF archives, while Khurana and Deshpande~\cite{KhuranaD16} propose indexes more generally for historical graph data. Bahri et al.~\cite{BahriLA18} explore the benefits of using Apache Spark to manage RDF archives in a distributed setting. In order to evaluate these emerging techniques, a number of benchmarks have been proposed, including the BEnchmark of RDF
ARchives (BEAR)~\cite{FernandezUPK19}, and the Semantic Publishing Benchmark (SPB)~\cite{Papakonstantinou18}.

@@

We now reflect back on our research questions: (Q1) In terms of indexing space and time, incremental builds with an initial base version (\cbpd) are best. (Q2) In general the uncompressed IC representation (\ic) offers the best query runtimes, but interval TB (\tb) is quite competitive, and even outperforms IC for delta-version queries. (Q3) Rather than there being an overall winner, we note a time--space tradeoff where the larger representations offer faster queries and smaller representations slower queries. However, interval TB (\tb) arguably strikes the best balance if trying to optimise for both space and time.

For future work, it would be of interest to run experiments for other SPARQL engines and other RDF archive benchmarks. Also it would be interesting to run more diverse types of versioned queries, such as delta versions with larger gaps, queries returning versions as solutions, etc. There are also open questions relating to more optimal/concise query rewritings, and support for paths.\footnote{Online material: \url{https://github.com/HunterNacho/sparql-versioning}}

In conclusion: for hosting RDF archives, is SPARQL all you need? Our results tend to suggest that, with some caveats, an off-the-shelf SPARQL engine can be a solid option to meet the core needs, though of course specialised languages and systems could offer more features and consume less time and space.
